\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-08-25 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results
updated the signs in the potentials
updates

\usepackage{amsmath,graphicx,amsfonts,amssymb,amsthm,epsfig,mathrsfs,balance}
\usepackage{subcaption,tikz,color,multirow,algorithm,algpseudocode,algorithmicx}
\usepackage{cite,url,framed,bm,dsfont,lipsum}
\usepackage{cuted}
\setlength\stripsep{3pt plus 1pt minus 1pt}

@@

\author{Abhishek~Halder,~\IEEEmembership{Senior Member,~IEEE,} 
        Kenneth~F. Caluya,
        Pegah~Ojaghi,
        and~Xinbo~Geng% <-this % stops a space
\thanks{A. Halder and K.F. Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
\thanks{P. Ojaghi is with the Department of Computer Science and Engineering, University of California, Santa Cruz, CA 95064, USA, {\texttt{pojaghi@ucsc.edu}}.}%
\thanks{X. Geng is with the School of Electrical and Computer Engineering, Cornell University, Ithaca, NY 14850, USA, {\texttt{xg72@cornell.edu}}.} %
\thanks{This research was partially supported by the NSF award 1923278.}}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%

@@

We define the \emph{positive diagonal} matrices
\begin{align*}
\bm{M} &:= {\rm{diag}}\left(m_{1},\hdots,m_{n}\right), \nonumber\\
\bm{\Gamma} &:= {\rm{diag}}\left(\gamma_{1},\hdots,\gamma_{n}\right),\nonumber\\
\bm{\Sigma} &:= {\rm{diag}}\left(\sigma_{1},\hdots,\sigma_{n}\right),
%\label{PosDiagMatrices}	
\end{align*}
and rewrite (\ref{ItoSDEcomponentlevel}) as a mixed conservative-dissipative SDE in state vector $\bm{x} := (\bm{\theta},\bm{\omega})^{\top} \in \mathbb{T}^{n} \times \mathbb{R}^{n}$ as
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\theta}\\
{\mathrm{d}}\bm{\omega}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\omega}\\
-\bm{M}^{-1}\nabla_{\bm{\theta}}V(\bm{\theta}) -\bm{M}^{-1}\bm{\Gamma}\bm{\omega}  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{M}^{-1}\bm{\Sigma}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{ItoSDEvectorlevel}
\end{align}
where 
%$\odot$ denotes element-wise product, $\oslash$ denotes element-wise division, the , the rotor inertia vector $\bm{m} = (m_{1},\hdots,m_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the rotor damping coefficient vector $\bm{\gamma} = (\gamma_{1}, \hdots, \gamma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the diffusion coefficient vector $\bm{\sigma} = (\sigma_{1}, \hdots, \sigma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, and 
$\bm{w}\in\mathbb{R}^{n}$ is the standard vector Wiener process, $\mathbb{T}^{n}$ denotes the $n$-torus $[0,2\pi)^{n}$, and the potential function $V : \mathbb{T}^{n} \mapsto \mathbb{R}$ is given by
\begin{align}
V(\bm{\theta}) := \displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{i,j=1}^{n}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right).
\label{potential}	
\end{align}
%wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network topology. 
The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}. %The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{\theta}\|_{2}\rightarrow +\infty$.
%In the development below, an important role will be played by the ``Hamiltonian-like" function
%\begin{align}
%H(\bm{x})\equiv H(\bm{\theta},\bm{\omega}) := \frac{1}{2}\bm{\omega}^{\top}{\rm{diag}}\left(\bm{\gamma}\oslash\bm{m}\right)\bm{\omega} \: + \: V(\bm{\theta}).
%\label{hamiltonian}	
%\end{align}

@@

\subsection{From Anisotropic to Isotropic Degenerate Diffusion}\label{subsec:FromAnisoToIso}
Consider the $2n\times 2n$ matrix
\begin{align}
\bm{\Psi}:= \bm{I}_{2} \otimes \left(\bm{M}\bm{\Sigma}^{-1}\right), \label{DefPsi}	
\end{align}
and define the invertible linear map
\begin{align}
\begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix} \mapsto \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix} := \bm{\Psi} \begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix}. \label{ChangeOfVar}
\end{align}
Applying It\^{o}'s lemma \cite[Ch. 4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{ItoSDEvectorlevel}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\xi}\\
{\mathrm{d}}\bm{\eta}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\eta}\\
-\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{I}_{n}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{XiEtaVectorSDE}	
\end{align}
where the potentials
{\small{\begin{subequations}
\begin{align}
U(\bm{\xi}) &:= \!\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} + \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
\end{align}
\label{NewPotentials}	
\end{subequations}}}

@@

@@

We randomly generated the noise coefficients $\sigma_{i}\in [1,5]$ as
\[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]

@@

With step size $h=10^{-3}$, we discretized time $t\in[0,1]$. Using the algorithm detailed in Sec. \ref{sec:ProxAlgo}, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$. In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $L=300$.

Fig. \ref{fig:IEEE14marginalsCase0} and Fig. \ref{fig:IEEE14marginalsCase1} show the time snapshots of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs for \textbf{Case I} and \textbf{Case II} respectively, associated with the generator $i=1,\hdots, n=5$. These marginals are computed using the corresponding weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$. %Notice that in Fig. \ref{fig:IEEE14marginals}, the horizontal axes of the subfigures have periodic boundary conditions, i.e., $(\theta_{i},\omega_{i})\in\mathbb{T}\times\mathbb{R}$ (cylinder). 
Fig. \ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.  \ref{sec:ProxAlgo}.

>>>>

\usepackage{amsmath,graphicx,amsfonts,amssymb,amsthm,epsfig,mathrsfs,balance}
\usepackage{subfigure,tikz,color,multirow,algorithm,algpseudocode,algorithmicx}
\usepackage{cite,url,framed,bm,dsfont,lipsum}
\usepackage{cuted}
\setlength\stripsep{3pt plus 1pt minus 1pt}

@@

\author{Abhishek~Halder,~\IEEEmembership{Senior Member,~IEEE,} 
        Kenneth~F. Caluya,
        Pegah~Ojaghi,
        and~Xinbo~Geng,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
\thanks{A. Halder and K.F. Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
\thanks{P. Ojaghi is with the Department of Computer Science and Engineering, University of California, Santa Cruz, CA 95064, USA, {\texttt{pojaghi@ucsc.edu}}.}%
\thanks{X. Geng is with the School of Electrical and Computer Engineering, Cornell University, Ithaca, NY 14850, USA, {\texttt{xg72@cornell.edu}}.} %
\thanks{This research was partially supported by the NSF award 1923278.}}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%

@@

We define the \emph{positive diagonal} matrices
\begin{align*}
\bm{M} &:= {\rm{diag}}\left(m_{1},\hdots,m_{n}\right), \nonumber\\
\bm{\Gamma} &:= {\rm{diag}}\left(\gamma_{1},\hdots,\gamma_{n}\right),\nonumber\\
\bm{\Sigma} &:= {\rm{diag}}\left(\sigma_{1},\hdots,\sigma_{n}\right),
%\label{PosDiagMatrices}	
\end{align*}
and rewrite (\ref{ItoSDEcomponentlevel}) as a mixed conservative-dissipative SDE in state vector $\bm{x} := (\bm{\theta},\bm{\omega})^{\top} \in \mathbb{T}^{n} \times \mathbb{R}^{n}$ as
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\theta}\\
{\mathrm{d}}\bm{\omega}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\omega}\\
-\bm{M}^{-1}\nabla_{\bm{\theta}}V(\bm{\theta}) -\bm{M}^{-1}\bm{\Gamma}\bm{\omega}  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{M}^{-1}\bm{\Sigma}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{ItoSDEvectorlevel}
\end{align}
where 
%$\odot$ denotes element-wise product, $\oslash$ denotes element-wise division, the , the rotor inertia vector $\bm{m} = (m_{1},\hdots,m_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the rotor damping coefficient vector $\bm{\gamma} = (\gamma_{1}, \hdots, \gamma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the diffusion coefficient vector $\bm{\sigma} = (\sigma_{1}, \hdots, \sigma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, and 
$\bm{w}\in\mathbb{R}^{n}$ is the standard vector Wiener process, $\mathbb{T}^{n}$ denotes the $n$-torus $[0,2\pi)^{n}$, and the potential function $V : \mathbb{T}^{n} \mapsto \mathbb{R}$ is given by
\begin{align}
V(\bm{\theta}) := -\displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{i,j=1}^{n}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right).
\label{potential}	
\end{align}
%wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network topology. 
The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}. %The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{\theta}\|_{2}\rightarrow +\infty$.
%In the development below, an important role will be played by the ``Hamiltonian-like" function
%\begin{align}
%H(\bm{x})\equiv H(\bm{\theta},\bm{\omega}) := \frac{1}{2}\bm{\omega}^{\top}{\rm{diag}}\left(\bm{\gamma}\oslash\bm{m}\right)\bm{\omega} \: + \: V(\bm{\theta}).
%\label{hamiltonian}	
%\end{align}

@@

\subsection{From Anisotropic to Isotropic Degenerate Diffusion}\label{subsec:FromAnisoToIso}
Consider the $2n\times 2n$ matrix
\begin{align}
\bm{\Psi}:= \bm{I}_{2} \otimes \left(\bm{M}\bm{\Sigma}^{-1}\right), \label{DefPsi}	
\end{align}
and define the invertible linear map
\begin{align}
\begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix} \mapsto \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix} := \bm{\Psi} \begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix}. \label{ChangeOfVar}
\end{align}
Applying It\^{o}'s lemma \cite[Ch. 4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{ItoSDEvectorlevel}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\xi}\\
{\mathrm{d}}\bm{\eta}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\eta}\\
-\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{I}_{n}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{XiEtaVectorSDE}	
\end{align}
where the potentials
{\small{\begin{subequations}
\begin{align}
U(\bm{\xi}) &:= \!-\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
\end{align}
\label{NewPotentials}	
\end{subequations}}}

@@

\begin{figure*}[ht!]
   \subfigure[]{\includegraphics[width=0.322\textwidth]{Case0_OmegaMarginals.png}}
   \hspace{\fill}
   \subfigure[]{\includegraphics[width=0.322\textwidth]{Case1_OmegaMarginals.png}}
   \hspace{\fill}
   \subfigure[]{\includegraphics[width=0.309\textwidth]{Case0Case1Combined_Gen4_OmegaMarginals.png}}
\caption{\small{For $i=1,\hdots,5$, the $\omega_{i}$ marginals (a) for Case I, and (b) for Case II, resulting from the IEEE 14 bus simulation described in Sec. \ref{subsec:NumSimIEEEBus}. (c) Comparison of the $\omega_{4}$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios.}}
\vspace*{-0.1in}
\label{fig:IEEE14omegamarginalsCase12}
\end{figure*}

@@

Our numerical simulations for the IEEE 14 bus system consider the following three cases:

\noindent\textbf{Case I.} Nominal case where ...

\noindent\textbf{Case II.} post-contingency simulation after the line 13 fails.

\noindent\textbf{Case III.}	load data are time-varying ...

For all the three cases, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
\[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]

@@

With step size $h=10^{-3}$, we discretized time $t\in[0,1]$. Using the algorithm detailed in Sec. \ref{sec:ProxAlgo}, for all the three cases, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$. In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $L=300$.

Fig. \ref{fig:IEEE14marginalsCase0} and Fig. \ref{fig:IEEE14marginalsCase1} show the time snapshots of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs for \textbf{Case I} and \textbf{Case II} respectively, associated with the generator $i=1,\hdots, n=5$. These marginals are computed using the corresponding weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.

 %Notice that in Fig. \ref{fig:IEEE14marginals}, the horizontal axes of the subfigures have periodic boundary conditions, i.e., $(\theta_{i},\omega_{i})\in\mathbb{T}\times\mathbb{R}$ (cylinder).

Fig. \ref{fig:IEEE14omegamarginalsCase12} shows the corresponding univariate $\omega_{i}$-marginal PDFs. In particular, Fig. \ref{fig:IEEE14omegamarginalsCase12}(c) helps compare the $\omega_{4}$ marginal evolution for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios. Since Case II considers the failure of line 13, we expect from Fig. \ref{fig:Motivating} that the transient effects of this fault will be prominent for bus 6. Indeed, Fig. \ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both Case I and II, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger dispersion.

Fig. \ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.  \ref{sec:ProxAlgo}.
